---
title: "Statistical Tests Part1"
date: "7/14/2020"
output: html_document
---

### Welcome to this week's SummeR of R session, our first of two sessions learning about statistical models/tests in R!  

Dr. Margarita Corral  

mcorral@brandeis.edu  

[Make an appointment with Margarita!](https://calendar.library.brandeis.edu/appointment/8518)  
  
  
Dr. Claire Pontbriand 

cpontbriand@brandeis.edu  

[Make an appointment with Claire!](https://calendar.library.brandeis.edu/appointments/scidata)


### Today we'll cover:
- Exploring and summarizing data
- Correlation and Chi-Squared test
- T-test and ANOVAS
- Checking assumptions


### Set up
To get started let's load or install the library's we'll be using. If this is your first time using one of the packages uncomment and run the appropriate install.package('package')
```{r}
#install.packages('tidyverse')
library(tidyverse)

#install.packages('car')
library(car)

#install.packages('broom')
library(broom)

#install.packages('dplyr')
library(dplyr)

#install.packages('ggplot2')
library(ggplot2)
```

We're going to be analyzing penguins! See https://allisonhorst.github.io/palmerpenguins/

```{r}
#install.packages("remotes")
remotes::install_github("allisonhorst/palmerpenguins")
library(palmerpenguins)
data(package = 'palmerpenguins')
```


## Let's start by exploring the data set
```{r}
View(penguins)
```
We can check the structure...
```{r}
str(penguins)
``` 
... and see the first six rows
```{r}
head(penguins)
```

I want to see how many penguins I have
```{r}
species_count <- count(penguins, species)
View(species_count)
```

I can get the same table using piping. Pipes express a sequence of multiple operations in a clear way.  The tidyverse loads %>% for you automatically
```{r}
penguins %>% 
  count(species)
```

I want to see summary statistics for each species
of penguin
```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), mean, na.rm = TRUE))
```

Is there a correlation between Flipper Lenght and Body Mass. Let's create a scatterplot first
```{r}
correlation_graph <- ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + geom_point() + geom_smooth(method = lm)
correlation_graph
```

What's the correlation coefficient?
```{r}
cor.test(penguins$flipper_length_mm, penguins$body_mass_g)
```

Now, I want to see if there is relationship between species and island. As both variables are categorical,we need to run a chi-squared test
```{r}
ggplot(penguins, aes(x = species, fill = island)) + geom_bar()
```

```{r}
chisq <- chisq.test(penguins$species, penguins$island)
chisq
```

Is there a difference between males and females in their body mass?
```{r}

ggplot(penguins)+
  geom_boxplot(aes(x=sex, y=body_mass_g, fill=sex))+
  theme_classic()+
  ylab("Body mass")+
  xlab('')+
  theme(legend.position = 'none')
```



# We can check to see if the two samples have equal variance by performing a leveneTest, inside the functions we put a formula of the structure 'continuous_variable ~ grouping'.
```{r}
leveneTest(body_mass_g~sex, data=penguins)
```


#Levene's test has a null hypothesis that the variances are equal,based on the results of our test we reject the null hypothesis  


# We'll use the *t.test()* to compare the two means, the default for the function is to assume variances are not equal, and it performs a Welch Two Sample t-test. If we have equal variance we have to set the argument var.equal=TRUE which would run a Two-sample t-test.  
```{r}
t.test(penguins$body_mass_g ~ penguins$sex)
```


## Comparing means from multiple groups

I would like to see if there are differences in the bill length between species
I am going to start with a visualization
```{r}
ggplot(penguins, aes(x = species, y = bill_length_mm)) + geom_boxplot() 
```

We can use the *aov()* function to run our anova.  We'll check our model assumptions after we fit the model. You want to assign the model fit to a variable name because we'll use it to get the statistics and check assumptions.
```{r}
anova<-aov(bill_length_mm~species, data=penguins)
```

```{r}
summary(anova)
```
Post-Hoc analysis to see where the differences are
```{r}
Tukey_test<- TukeyHSD(anova)
```



#This is kind of hard to work. Luckily the tidyverse has a package called broom (we loaded this in earlier) and broom has a function called *tidy()*
```{r}
Tukeys<-TukeyHSD(anova)%>%tidy()
```



#Let's highlight significant differences
```{r}
Tukeys<-Tukeys %>% mutate(sig = case_when(adj.p.value < .05~ '*', TRUE ~''))
Tukeys
```


#We can also get a plot
```{r}
plot(Tukey_test)
```


Checking the homogeneity of variance assumption. 
```{r}
leveneTest(bill_length_mm ~ species, data=penguins)
```
Checking the normality assumption 
```{r}
plot(anova, 2)
```
or we can use the Shapiro_Wilk test
```{r}
anova_residuals <- residuals(object = anova )
shapiro.test(x=anova_residuals)
```

The non-parametric alternative to one-way ANOVA is Kruskal-Wallis rank sum test, which can be used when ANNOVA assumptions are not met.

```{r}
kruskal.test(bill_length_mm~species, data=penguins)
```



### Good resources to check out for more variations/details:
* [R cookbook](https://search.library.brandeis.edu/permalink/f/1skfba6/TN_sbo_s9780596809287) by Paul Teetor
* [Learning Statistics with R](https://learningstatisticswithr.com/book/) by Danielle Navarro

### What to do now?
Practice wrangling and running stats on the titanic data set. Some questions you could answer:
  - Is there a significant difference in age of the people that survived vs. not? 
  - Is there a significant difference in age across passenger classes?
```{r}
#Load this package
#install.packages('titanic')
#library(titanic)
#The package has a dataframe called Titanic that is now available for you to access 
#View(Titanic)
```



